---
title: "CoV19"
author: Eli Holmes
output: github_document
---

This is a package for exploring some of the CoV-19 data. The world data is downloaded from JHU. The US data is downloaded from JHU and CovidTracking. The Italy data is downloaded from the Italian CDC (Protezione Civile) 

Sources:

* https://github.com/CSSEGISandData/COVID-19
* https://covidtracking.com/api/ 
* https://github.com/pcm-dpc/COVID-19

To install the package

```
library(devtools)
install_github(eeholmes/CoV19)
```

To get the data, type any of these on command line. After loading.
```
states
italy
world
```
Use `head()` to look at it. Should be pretty self-evident what it is.

# Plots

There aren't many. Yet. The basic one I am working on is new cases versus cases after a lockdown.

```{r, message=FALSE, warning=FALSE}
library(CoV19)
plot1("states", "WA", lockdown="2020-03-16")
```

```{r, message=FALSE, warning=FALSE}
plot1("italy", "Lombardia", lockdown="2020-03-11")
```

Something is odd with the Hubei data in the JHU data set.

```{r, message=FALSE, warning=FALSE}
plot1("world", "Hubei", lockdown="2020-01-23")
```
