---
title: "CoV19"
author: Eli Holmes
output: github_document
---

This is a package for exploring some of the CoV-19 data. The world data is downloaded from JHU. The US data is downloaded from JHU and CovidTracking. The Italy data is downloaded from the Italian CDC (Protezione Civile) 

Sources:

* https://github.com/CSSEGISandData/COVID-19
* https://covidtracking.com/api/ 
* https://github.com/pcm-dpc/COVID-19

To install the package

```
library(devtools)
install_github("eeholmes/CoV19"")
```

To get the data, type any of these on command line. After loading.
```
states
italy
world
```
Use `head()` to look at it. Should be pretty self-evident what it is.

# Functions

* `regions()` will show you the regions in each data set. 
* `plot2()` a basic plot of total, hospitalized, and deaths. First argument is a data object (states, italy, world or subset of these) and second argument is a region (can be partial). Use ?plot2 to read the help page.

# Plots

```{r, message=FALSE, warning=FALSE}
library(CoV19)
plot2(italy, "Lomb", decorate=TRUE)
```

```{r, message=FALSE, warning=FALSE}
plot2(states, c("WA","CA"))
```

```{r, message=FALSE, warning=FALSE}
plot2(states, c("NY","NJ","PA","MA"))
```

Something is odd with the Hubei data in the JHU data set.

```{r, message=FALSE, warning=FALSE}
plot2(world, "Hubei China", decorate=TRUE)
```
